As-you-go" instead of "after-the-fact": A network approach to scholarly communication and evaluation

### 2018-03-30 docs/manuscript.Rmd

  We propose an alternative design for scholarly communication based on modular research outputs with direct links between subsequent modules, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, modular communication was proposed two decades ago [@Kircz1998;@VandeSompel2004;@10.7717/peerj-cs.78;@10.3233/ISU-2010-0613;@10.1163/095796511X560006]. These modules could be similar to sections of a research paper, but extend to modular research outputs such as software or materials. We propose to implement this modular communication on an "as-you-go" basis and include direct links to indicate provenance. This respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

@@

  Networks indicators could be used to evaluate the network as it exists now or how it developed in the past (i.e., backward-looking evaluation). For example, in-degree centrality could be used to identify highly interconnected modules of information. This measure indicates how many child modules are spawned by a parent module and indicates how much new work a researcher's output stimulates (e.g., module04 in Table 1 would have an in-degree centrality of three). To contextualize this, a data module could spawn four results sections, hence has an in-degree centrality of four. This measure would look only at one-generation of child modules, but other measures extend this to incorporate multiple generations of child modules. Katz centrality extends this and computes the centrality over $N$ generations of child modules [pp. 206-210; @Wasserman1994] whereas traditional in-degree centrality calculates centrality for $N=1$ generations. For example, two data modules that each spawn five results modules would have the same in-degree centrality, but could have different Katz centrality if only one of those two networks has a third-generation of modules included. If multi-generation indicators are relevant, Katz centrality measures could provide operationalizations of such measures.

  Another set of network indicators could be used to evaluate how the network would change when new modules are added in the future (i.e., forward-looking evaluation). For example, a researcher who is looking for ways to increase the density in their own network, could ask the question "If I would add one module that has $N$ parents, which addition would increase the density the most?" Subsequently, the researcher could inspect the identified connections for inspiration and feasibility. Complexity of the new module could be increased by increasing the number of parent modules to connect ($N$ in the question; e.g., five instead of two). Potentially, this could facilitate creative thinking, where $N$ is gradually increased over time to increase the complexity of the issue from a network perspective.

  The indicators we highlighted here are simple proposals. Indicators from network analysis and graph theory could be applied to the study of knowledge development when a network structure is available and we hope to see suggestions to answer questions about the network. These kinds of analyses are already done within citation networks [e.g., @10.1126/science.aao0185] and authorship networks [e.g., @10.1371/journal.pntd.0000501], but we cannot do so with the provenance or planning of knowledge generation in the current scholarly communication system.

@@

  To encourage culture- and behavioral change, "after-the-fact" papers and modular "as-you-go" outputs could co-exist (initially) and would not require researchers to make a zero-sum decision. Copyright is often transferred to publishers upon publication (resulting in pay-to-access), but only after a legal contract is signed. Hence, preprints cannot be legally restricted by publishers when they precede a copyright transfer agreement. However, preprints face institutional and social opposition [@Kaiser2017], where preprinting could exclude a manuscript for publication depending on editorial policies or due to fears of non-publication or scooping (itself a result of hypercompetition). In recent years, preprints have become more widely accepted and less likely to exclude manuscript publication (e.g., Science accepts preprinted manuscripts) [@Berg2017]. Similarly, sharing modular "as-you-go" outputs would not be legally restricted by publishers and could ride the wave of preprint acceptance, but might also face institutional or social counterchange similar to the history of preprints. Researchers could communicate "as-they-go" and compile "after-the-fact" papers, facilitating co-existence and minimizing negative effects on career opportunities. Additionally, “as-you-go” modules could be used in scholarly fields where the provenance of information is important to findings.

@@

  The current scholarly communication system based on research papers is “after-the-fact” and can be supplemented by an “as-you-go” based communication system. By doing so, the functions of a scholarly communication system can be interpreted more widely, making registration complete, certification part of the process instead of just the judgment of peers, access to everything for everyone based on peer-to-peer protocols, simplify archival, and facilitate incentive structures that could align researcher’s interests with that of scholarly research.

>>>>

  We propose an alternative design for scholarly communication based on modular research outputs with direct links between subsequent modules, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, modular communication was proposed two decades ago [@Kircz1998;@VandeSompel2004;@10.7717/peerj-cs.78;@10.3233/ISU-2010-0613;@10.1163/095796511X560006;Nielsen2012]. These modules could be similar to sections of a research paper, but extend to modular research outputs such as software or materials. We propose to implement this modular communication on an "as-you-go" basis and include direct links to indicate provenance. This respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

@@

  Networks indicators could be used to evaluate the network as it exists now or how it developed in the past (i.e., backward-looking evaluation). For example, in-degree centrality could be used to identify highly interconnected modules of information. This measure indicates how many child modules are spawned by a parent module and indicates how much new work a researcher's output stimulates (e.g., module04 in Table 1 would have an in-degree centrality of three). To contextualize this, a data module could spawn four results modules, hence has an in-degree centrality of four. This measure would look only at one-generation of child modules, but other measures extend this to incorporate multiple generations of child modules. Katz centrality extends this and computes the centrality over $N$ generations of child modules [pp. 206-210; @Wasserman1994] whereas traditional in-degree centrality calculates centrality for $N=1$ generations. For example, two data modules that each spawn five results modules would have the same in-degree centrality, but could have different Katz centrality if only one of those two networks has a third-generation of modules included. If multi-generation indicators are relevant, Katz centrality measures could provide operationalizations of such measures.

  Another set of network indicators could be used to evaluate how the network would change when new modules are added in the future (i.e., forward-looking evaluation). For example, a researcher who is looking for ways to increase the density in their own network, could ask the question "If I would add one module that has $k$ parents, which addition would increase the density the most?" Subsequently, the researcher could inspect the identified connections for inspiration and feasibility. Complexity of the new module could be increased by increasing the number of parent modules to connect ($k$ in the question; e.g., five instead of two). Potentially, this could facilitate creative thinking, where $k$ is gradually increased over time to increase the complexity of the issue from a network perspective.

  The indicators we highlighted here are simple proposals. Other indicators from network analysis and graph theory could be applied to the study of knowledge development when a network structure is available and we hope to see suggestions to answer questions about the network. These kinds of analyses are already done within citation networks (e.g., [@10.1126/science.aao0185]) and authorship networks (e.g., [@10.1371/journal.pntd.0000501]), but we cannot do so with the provenance or planning of knowledge generation in the current scholarly communication system.

@@

  To encourage culture- and behavioral change, "after-the-fact" papers and modular "as-you-go" outputs could co-exist (initially) and would not require researchers to make a zero-sum decision. Copyright is often transferred to publishers upon publication (resulting in pay-to-access), but only after a legal contract is signed. Hence, preprints cannot be legally restricted by publishers when they precede a copyright transfer agreement. However, preprints face institutional and social opposition [@Kaiser2017], where preprinting could exclude a manuscript for publication depending on editorial policies or due to fears of non-publication or scooping (itself a result of hypercompetition). In recent years, preprints have become more widely accepted and less likely to exclude manuscript publication (e.g., Science accepts preprinted manuscripts) [@Berg2017]. Similarly, sharing modular "as-you-go" outputs would not be legally restricted by publishers and could ride the wave of preprint acceptance, but might also face institutional or social counterchange similar to the history of preprints. Researchers could communicate "as-they-go" and compile "after-the-fact" papers, facilitating co-existence and minimizing negative effects on career opportunities. Additionally, "as-you-go" modules could be used in scholarly fields where the provenance of information is important to findings.

@@

  The current scholarly communication system based on research papers is "after-the-fact" and can be supplemented by an "as-you-go" based communication system. By doing so, the functions of a scholarly communication system can be interpreted more widely, making registration complete, certification part of the process instead of just the judgment of peers, access to everything for everyone based on peer-to-peer protocols, simplify archival, and facilitate incentive structures that could align researcher’s interests with that of scholarly research.
